Financial markets for reaching ``green" targets: Cap-and-trade strategies in the Chilean electricity sector

### 2020-04-11 draft_paper/draft_v1.tex
typos y parametrización modelo

Several studies have been published about the the impact of carbon pricing and the carbon tax in Chile. There is a broad agreement that the \$5 per tCO$_2$ tax is not an efficient incentive to reduce emissions. Mardones and Flores \cite{mardones2017evaluation} performed an economic analysis to find the best emission tax, by estimating the fuel consumption and the carbon emissions. They concluded that the tax fails to generate a change in fuels, and thus, mitigate emissions, even for different costs scenarios. They estimated that a more effective tax is between \$13 - \$50 per  tCO$_2$. In addition, the economic and environmental effects of a carbon tax in Latin American countries with such implemented policy (Chile and M\'exico) and one country discussing it (Brazil) was studied by \cite{mardones2018economic}. Authors concluded that in all three countries the adopted tax is not enough to achieved the commitments made by each country in the Paris Agreement, particularly in the case of Chile. This issue is mentioned in the roadmap to 2050 document \cite{HojadeRuta}, noting that by 2030, the tax will increase to \$25 per tCO$_2$, which still is insufficient according to the studies cited above. Vera and Sauma \cite{vera2015does} also found that the Chilean carbon tax does not comply its goal, and, furthermore, they suggested that energy efficiency measures, such as introduction of flexible tariff schemes, more efficient heating system or even replacement of light and appliances, may reduce more carbon emissions than the tax, without increasing the marginal cost.

@@

Our model assumes two basic stages: In the first stage (period $t=0$), electric producers decide their generation, capacity investment and carbon allowance allocation to satisfy an exogenous demand $D(0)\in\mathbb{R}_+$. In the second stage ($t\in T:=\{1,...\bar{t}\}$), uncertainty is revealed in period $t=1$. Uncertainty is represented by a state of nature $\omega\in\Omega:=\{1,..,K\}$. Thus, for a given state of nature $\omega\in\Omega$, the pair $(t,\omega)$ represents the demand in period $t$ when the state $\omega$ is reached. For each pair $(t,\omega)$, given an initial allowance allocation in period $t=0$ and a secondary CO$_2e$ permit trading in $t=1$, electric generators engage in a spot market and decide their generation and new capacity investment such that the exogenous stochastic demand level $D(t,\omega)\in\mathbb{R}_+^{T\times\Omega}$ is satisfied. The demand for all periods is denoted by $D=\left(D(0),(D(t,\omega))_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$.

@@

\smallskip
The complete nomenclature of sets, parameters and variables are summarized in Table~\ref{tab:nom}. In the following paragraphs we describe the optimization problems for each agent in this market, and we define the existence of equilibrium for the problem.

@@

Defining $T_0:=\{0\}\cup T$, the optimization problem of producer $i$ is given by choosing $(x_i,Q_i, A_i,P_i,V_i)\in\mathbb{X}:=(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}) \times(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}) \times \mathbb{R}_+\times\mathbb{R}_+^{\Omega}\times\mathbb{R}_+^{\Omega}$ given prices $(\pi^d$, $\pi^a$ , $\pi^v)\in\Pi:=(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega})\times\mathbb{R}_+\times\mathbb{R}^{\Omega}_+$ and parameters $\left((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,\bar{Q}_i, RP_i , \varepsilon_i\right)\in \Xi:=\mathbb{R}_+^2\times\mathbb{R}_+^7$, $\tau\in \mathbb{R}_+$ and probability $(Pr(\omega))_{\omega\in\Omega}\in\Delta:=\left\{\left(Pr(\omega)\right)_{\omega\in\Omega}\in[0,1]^K:\sum_{\omega\in\Omega}Pr(\omega)=1\right\}$ as a solution of

@@

\subsection{Auctioneer's problem:}\label{sec:auctioneer}
 The auctioneer chooses the total number allowances available in the market $\theta\in \mathbb{R}_+$ and sells them at a price $\pi^a$. The price is set such that it clears the CO$_2e$ permit market among generators and the auctioneer.  The carbon budget (maximum level of CO$_2e$ emissions allowed in the second stage $T:=\{1,...\bar{t}\}$) is denoted by  $CAP\in\mathbb{R}_+$ and is drawn from a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$. Hence, the variable $\theta$ must not probabilistically surpass the CO$_2e$ budget for emissions, as shown in Equation~\ref{eq:probcap}.

@@

 The objective of the auctioneer is to maximize its benefit. Note that benefit is defined as the revenue obtained by selling allowances while considering negative outcomes modeled by $\mathcal{F}(\theta)$. For instance, $\mathcal{F}(\theta)$ can represent the cost of an additional tonne of CO$_2$ imposed to the society\footnote{The Chilean Ministry of Social Development estimated the cost of carbon for Chile to be equivalent to USD \$32 \cite{ChileSCC}.}, known as the Social Cost of Carbon (SCC)~\cite{feijoo2014design}. Thus, the optimization problem for the auctioneer is given by choosing $\theta$ given price $\pi^a$ and parameters $\sigma$, $\mu$ and $\epsilon$ as a solution of:

@@

%\subsection{Market clearing conditions:}

\begin{definition}
An equilibrium among production and capacity investment models and the auctioneer is a vector of prices and production decisions $$\left((\pi^{d*},\pi^{a*},\pi^{v*}),(x_i^*,Q_i^*,A_i^*,P_i^*,V_i^*)_{i\in\{1,...,N\}},\theta^*\right)\in\Pi\times\mathbb{X}^N \times \mathbb{R}_+$$ such that:

>>>>

Several studies have been published about the the impact of carbon pricing and the carbon tax in Chile. There is a broad agreement that the \$5 per tCO$_2$ tax is not an efficient incentive to reduce emissions. Mardones and Flores \cite{mardones2017evaluation} performed an economic analysis to find the best emission tax, by estimating the fuel consumption and the carbon emissions. They concluded that the tax fails to generate a change in fuels, and thus, mitigate emissions, even for different costs scenarios. They estimated that a more effective tax is between \$13 - \$50 per  tCO$_2$. In addition, the economic and environmental effects of a carbon tax in Latin American countries with such implemented policy (Chile and M\'exico) and one country discussing it (Brazil) was studied by \cite{mardones2018economic}. Authors concluded that in all three countries the adopted tax is not enough to achieve the commitments made by each country in the Paris Agreement, particularly in the case of Chile. This issue is mentioned in the roadmap to 2050 document \cite{HojadeRuta}, noting that by 2030, the tax will increase to \$25 per tCO$_2$, which still is insufficient according to the studies cited above. Vera and Sauma \cite{vera2015does} also found that the Chilean carbon tax does not comply its goal, and, furthermore, they suggested that energy efficiency measures, such as introduction of flexible tariff schemes, more efficient heating system or even replacement of light and appliances, may reduce more carbon emissions than the tax, without increasing the marginal cost.

@@

Our model assumes two basic stages: In the first stage (period $t=0$), electric producers decide their generation, capacity investment and carbon allowance allocation to satisfy an exogenous demand $D(0)\in\mathbb{R}_+$. In the second stage ($t\in T:=\{1,...\bar{t}\}$), uncertainty is revealed in period $t=1$. Uncertainty is represented by a state of nature $\omega\in\Omega:=\{1,..,K\}$. Thus, for a given state of nature $\omega\in\Omega$, the pair $(t,\omega)$ represents the demand in period $t>1$ when the state $\omega$ is reached. For each pair $(t,\omega)$, given an initial allowance allocation in period $t=0$ and a secondary CO$_2e$ permit trading in $t=1$, electric generators engage in a spot market and decide their generation and new capacity investment such that the exogenous stochastic demand level $D(t,\omega)\in\mathbb{R}_+^{T\times\Omega}$ is satisfied. The demand for all periods is denoted by $D=\left(D(0),(D(t,\omega))_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$.

@@

\smallskip
The complete nomenclature of sets, parameters and variables are summarized in Table~\ref{tab:nom}. In the following paragraphs we describe the optimization problems for each agent in this market, and we define the equilibrium concept for the configuration.

@@

Defining $T_0:=\{0\}\cup T$, the optimization problem of producer $i$ is given by choosing \linebreak$(x_i,Q_i, A_i,P_i,V_i)\in\mathbb{X}:=\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right) \times\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right) \times \mathbb{R}_+\times\mathbb{R}_+^{\Omega}\times\mathbb{R}_+^{\Omega}$ given prices \linebreak$(\pi^d$, $\pi^a$ , $\pi^v)\in\Pi:=\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right)\times\mathbb{R}_+\times\mathbb{R}^{\Omega}_+$ and parameters \linebreak$\left((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,\bar{Q}_i, RP_i , \varepsilon_i\right)\in \Xi:=\mathbb{R}_+^2\times\mathbb{R}_+^7$, $\tau\in \mathbb{R}_+$ and probability $(Pr(\omega))_{\omega\in\Omega}\in\Delta:=\left\{\left(Pr(\omega)\right)_{\omega\in\Omega}\in[0,1]^K:\sum_{\omega\in\Omega}Pr(\omega)=1\right\}$ as a solution of

@@

\subsection{Auctioneer's problem:}\label{sec:auctioneer}
 The auctioneer chooses the total number allowances available in the market $\theta\in \mathbb{R}_+$ and sells them at a price $\pi^a$ in stage $t=0$. The price is set such that it clears the CO$_2e$ permit market among generators and the auctioneer.  The carbon budget (maximum level of CO$_2e$ emissions allowed in the second stage $t\in T$) is denoted by  $CAP\in\mathbb{R}_+$ and is drawn from a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$. Hence, the variable $\theta$ must not probabilistically surpass the CO$_2e$ budget for emissions, as shown in Equation~\ref{eq:probcap}.

@@

 The objective of the auctioneer is to maximize its benefit. Note that benefit is defined as the revenue obtained by selling allowances while considering negative outcomes modeled by $\mathcal{F}(\theta)$. For instance, $\mathcal{F}(\theta)$ can represent the cost of an additional tonne of CO$_2$ imposed to the society, known as the Social Cost of Carbon (SCC)~\cite{feijoo2014design}.\footnote{The Chilean Ministry of Social Development estimated the cost of carbon for Chile to be equivalent to USD \$32 \cite{ChileSCC}.} Thus, the optimization problem for the auctioneer is given by choosing $\theta$ given price $\pi^a$ and parameters $\sigma$, $\mu$ and $\epsilon$ as a solution of:

@@

%\subsection{Market clearing conditions:}
The data that parameterize the capacity investment model is given by the tuple $$\left(\tau,(Pr(\omega))_{\omega\in\Omega},\left(CAP,\mu,\sigma,\epsilon\right),\left((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,\bar{Q}_i, RP_i , \varepsilon_i\right)_{i\in N}\right)\in \mathbb{R}_+\times \Delta\times\mathbb{R}_+^4\times\Xi^N,$$

where the equilibrium notion is given by the following definition.

\begin{definition}
An equilibrium in the capacity investment model is a vector of prices and production decisions $$\left((\pi^{d*},\pi^{a*},\pi^{v*}),(x_i^*,Q_i^*,A_i^*,P_i^*,V_i^*)_{i\in\{1,...,N\}},\theta^*\right)\in\Pi\times\mathbb{X}^N \times \mathbb{R}_+$$ such that:
